Regression Models Project
=========================

Author: Kevin Krenz
Date: November 22, 2014

## Executive summary

Fuel economy is of growing importance as the price of gasoline increases, as well as our knowledge of the impact of burning fossil fuels. One aspect of cars that could have a significant impact on fuel economoy is the type of transmission - automatic or manual. This brief report will analyze the impact of transmission type on fuel economy from a sample of 32 cars in a 1974 Motor Trends magazine.

I find that ***

```{r load, echo=FALSE}

## Load the data, make transmission type a factor
data(mtcars)
mtcars$am <- factor(mtcars$am, levels=c(0, 1), labels=c('Automatic', 'Manual'))
```



## The data

```{r means, echo=FALSE}
means <- round(tapply(mtcars$mpg, mtcars$am, mean), 2)
sds <- round(tapply(mtcars$mpg, mtcars$am, sd), 2)

n <- dim(mtcars[1])
n.auto <- sum(mtcars$am == 'Automatic')
n.man <- sum(mtcars$am == 'Manual')
```

Our dataset contains data on `r n` different cars. The variables include milers per gallon, number of cylinders, displacement, gross horsepower, rear axle ratio, weight (lb/1000), quarter mile time, V/S, transmission type, number of forward gears, and the number of carburetors. 

Highlighting our variables of interest, there are `r n.auto` cars with automatic transmissions. They have a mean MPG of `r means[1]` with a standard deviation of `r sds[1]`. There are `r n.man` cars with manual transmissions. They have a mean MPG of `r means[2]` with a standard deviation of `r sds[2]`.

On first glance, then, it appears that cars with a manual transmission are more fuel efficient. However, that may not be the cause; it is possible that the cars with a manual transmission happen to share some other characteristics.

## Models



```{r echo=FALSE}
model1 <- lm(mpg ~ am + wt, mtcars)
model2 <- lm(mpg ~ am + wt + drat, mtcars)
model3 <- lm(mpg ~ am + wt + drat, mtcars)
model4 <- lm(mpg ~ am + wt + drat + am*drat, mtcars)
```


## Model selection

To select a model, I will look at the:

- $r^2$ values to measure percent of variation explained, and
- dfbetas to consider the impact of each sample on the coefficients, and
- hat values to consider the influence of samples, and
- residual plots to analyze the appropriateness of fit.


## Conclusion


## Appendix 1: Figures

#### Figure 1
```{r}
par(mfrow=c(1,2))
with(mtcars, plot(wt, mpg, pch=19, col=am, xlab='Weight (1000s pounds)',
                  ylab='MPG'))
with(mtcars, plot(drat, mpg, pch=19, col=am, xlab='Rear axle ratio',
                  ylab='MPG'))
```
